A Gentle Introduction and Application of Feature-Based Clustering with Psychological Time Series

### 2023-07-29 latex-results/07ClusterEvaluationInterpretation.tex
feedback lfb

Finally, we can also assess the clusters across other individual
difference variables \citep[e.g.,][]{monden2022}. This out-of-feature
comparison allows us to check for data artifacts, as well as check
whether the developmental clusters are associated with important social
markers and individual differences. To illustrate artifact checks, we
added the number of ESM measurements into the comparison and find that
the participants in the deterioration cluster (cluster 2) on average
completed slightly more ESM surveys in general and reported on more
intergroup interactions in particular (see \(n\) in
\fgrref[B]{fig:clusterFeatVar}). In our data exclusion procedures, we
ensured that the general time frame and completion rates are similar for
all participants and indeed the numbers in ESM measurements generally
are largely similar (e.g., see \(n\) for well-being and outgroup
attitudes). However, the difference in the reported number of
interactions might indicate either a clustering artifact or a meaningful
difference. The higher average number of interactions in cluster 2
could, for example, indicate a clustering artifact if variances are more
inflated due to the larger samples \citep{kogan2006}. In our case this
seems less likely because one out of four variables did not differ in
terms of the MAD (see \fgrref{fig:cluster_comparison_mad}). At the same
time, however, the difference in the number of experienced interactions
might also indicate a meaningful difference, where the deteriorating
cluster (cluster 2) on average reported more outgroup interactions
(\(difference\) = 1.028, \(t\)(150.83) = 7.50, \(p\) \textless{} .001,
\(95%CI [0.76, 1.30]), but these interactions were less voluntary (
\)difference\$ = -1.04, \(t\)(108.89) = -7.71, \(p\) \textless{} .001,
\(95%CI [-1.31, -0.77]), less meaningful (
\)difference\$ = -1.003, \(t\)(136.40) = -7.16, \(p\) \textless{} .001,
\(95%CI [-1.28, -0.73]), and less positive (
\)difference\$ = -1.38, \(t\)(152.31) = -11.94, \(p\) \textless{} .001,
\$95\%CI {[}-1.61, -1.15{]}). Such a finding is in line with past
research highlighting the role of negative intergroup interactions in
explaining intergroup relations
\citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.

>>>>

Finally, we can also assess the clusters across other individual
difference variables \citep[e.g.,][]{monden2022}. This out-of-feature
comparison allows us to check for data artifacts, as well as check
whether the developmental clusters are associated with important social
markers and individual differences. To illustrate artifact checks, we
added the number of ESM measurements into the comparison and find that
the participants in the deterioration cluster (cluster 2) on average
completed slightly more ESM surveys in general and reported on more
intergroup interactions in particular (see \(n\) in
\fgrref[B]{fig:clusterFeatVar}). In our data exclusion procedures, we
ensured that the general time frame and completion rates are similar for
all participants and indeed the numbers in ESM measurements generally
are largely similar (e.g., see \(n\) for well-being and outgroup
attitudes). However, the difference in the reported number of
interactions might indicate either a clustering artifact or a meaningful
difference. The higher average number of interactions in cluster 2
could, for example, indicate a clustering artifact if variances are more
inflated due to the larger samples \citep{kogan2006}. In our case this
seems less likely because one out of four variables did not differ in
terms of the MAD (see \fgrref{fig:cluster_comparison_mad}). At the same
time, however, the difference in the number of experienced interactions
might also indicate a meaningful difference, where the deteriorating
cluster (cluster 2) on average reported more outgroup interactions
(\(difference\) = 1.03, \(t\)(150.83) = 7.50, \(p\) \textless{} .001,
\(95%CI
\) {[}0.76, 1.30{]}), but these interactions were less voluntary
(\(difference\) = -1.04, \(t\)(108.89) = -7.71, \(p\) \textless{} .001,
\(95%CI
\) {[}-1.31, -0.77{]}), less meaningful (\(difference\) = -1.00,
\(t\)(136.40) = -7.16, \(p\) \textless{} .001, \(95%CI
\) {[}-1.28, -0.73{]}), and less positive (\(difference\) = -1.38,
\(t\)(152.31) = -11.94, \(p\) \textless{} .001, \(95%CI
\) {[}-1.61, -1.15{]}). Such a finding is in line with past research
highlighting the role of negative intergroup interactions in explaining
intergroup relations \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.

### 2023-07-29 sections/07ClusterEvaluationInterpretation.Rmd
feedback lfb

  diff <- results_df %>% 
    filter(concept == concept_in, feature == feature_in) %>% 
    select(difference) %>% 
    pull %>% 
    format(round(., 2), nsmall=2)
  t <- results_df %>% 
    filter(concept == concept_in, feature == feature_in) %>% 
    select(t_value) %>% 
    pull %>% 
    round(., 2) %>% 
    format(., nsmall=2)
  df <- results_df %>% 
    filter(concept == concept_in, feature == feature_in) %>% 
    select(df) %>% 
    pull %>% 
    round(., 2) %>%
    format(., nsmall=2)
  p <- results_df %>% 
    filter(concept == concept_in, feature == feature_in) %>% 
    select(p_value) %>% 
    pull
  p <- ifelse(p < 0.001, "< .001", 
              paste0("= ", format(round(p, 3), nsmall=3)))
  lwr <- results_df %>% 
    filter(concept == concept_in, feature == feature_in) %>% 
    select(lwr) %>% 
    pull %>% 
    round(., 2) %>%
    format(., nsmall=2)
  upr <- results_df %>% 
    filter(concept == concept_in, feature == feature_in) %>% 
    select(upr) %>% 
    pull %>% 
    round(., 2) %>%
    format(., nsmall=2)

  paste("$difference$ = ", diff,
        ", $t$(", df, ") = ", t,
        ", $p$ ", p , 
        ", $95%CI [", lwr,
        ", ", upr, "]",
        sep = "")
}
```

>>>>

  diff <- results_df %>% 
    filter(concept == concept_in, feature == feature_in) %>% 
    select(difference) %>% 
    pull %>% 
    round(., 2) %>% 
    format(., nsmall=2)
  t <- results_df %>% 
    filter(concept == concept_in, feature == feature_in) %>% 
    select(t_value) %>% 
    pull %>% 
    round(., 2) %>% 
    format(., nsmall=2)
  df <- results_df %>% 
    filter(concept == concept_in, feature == feature_in) %>% 
    select(df) %>% 
    pull %>% 
    round(., 2) %>%
    format(., nsmall=2)
  p <- results_df %>% 
    filter(concept == concept_in, feature == feature_in) %>% 
    select(p_value) %>% 
    pull
  p <- ifelse(p < 0.001, "< .001", 
              paste0("= ", format(round(p, 3), nsmall=3)))
  lwr <- results_df %>% 
    filter(concept == concept_in, feature == feature_in) %>% 
    select(lwr) %>% 
    pull %>% 
    round(., 2) %>%
    format(., nsmall=2)
  upr <- results_df %>% 
    filter(concept == concept_in, feature == feature_in) %>% 
    select(upr) %>% 
    pull %>% 
    round(., 2) %>%
    format(., nsmall=2)

  paste("$difference$ = ", diff,
        ", $t$(", df, ") = ", t,
        ", $p$ ", p , 
        ", $95%CI$ [", lwr,
        ", ", upr, "]",
        sep = "")
}
```
